### 2025-09-10 inst/manuscript/manuscript.Rmd
some updates

---
documentclass: jss
author:
    # use this syntax to add text on several lines
    # To add another line, use \AND at the end of the previous one as above
    # use a different affiliation in adress field (differently formated here)
  - name: Michael Dumelle
    orcid: 0000-0002-3393-5529
    affiliation: |
      | United States
      | Environmental Protection Agency
    address: |
      | 200 SW 35th St
      | Corvallis, OR, 97330
    email: \email{Dumelle.Michael@epa.gov}
  - name: Jay M. Ver Hoef
    orcid: 0000-0003-4302-6895
    affiliation: |
      | Alaska Fisheries
      | Science Center
  - name: Matt Higham
    orcid: 0009-0006-4217-625X
    affiliation: |
      | St. Lawrence University
title:
  # If you use tex in the formatted title, also supply version without
  # For running headers, if needed
  formatted: "Spatial Generalized Linear Models in \\proglang{R} Using \\pkg{spmodel}"
  plain:     "Spatial Generalized Linear Models in R Using spmodel"
  short:     "Spatial Generalized Linear Models in \\proglang{R} Using \\pkg{spmodel}"
abstract: |
  Non-Gaussian data are common in practice and include binary, count, skewed, and proprtion data types. Often, non-Gaussian data are modeled using a generalized linear model (GLM). GLMs typically assume that observations are independent of one another. This is an impractical assumption for spatial data, as nearby observations tend to be more similar than distant ones. The \pkg{spmodel} package in \proglang{R} provides a suite of tools for fitting spatial generalized linear models (SPGLMs) to non-Gaussian data and making spatial predictions (i.e., Kriging). SPGLMs for point-referenced data (x- and y-coordinates) are fit using the \code{spglm()} function, while SPGLMs for areal (lattice, polygon) data are fit using the \code{spgautor()} function. Both \code{spglm()} and \code{spgautor()} maximize a novel Laplace likelihood which marginalizes over the model's fixed effects and latent mean while formally incorporating spatial covariance (i.e., spatial correlation). The inputs and outputs of \code{spglm()} and \code{spgautor()} closely resemble the \code{glm()} function from base \proglang{R}, easing the transition from GLMs to SPGLMs. \pkg{spmodel} provides and builds upon several commonly used helper functions for model building like \code{summary()}, \code{plot()}, \code{fitted()}, and \code{tidy()}, among others. Spatial predictions of the latent mean at unobserved locations are obtained using \code{predict()} or \code{augment()}. \pkg{spmodel} accommodates myriad advanced modeling features like geometric anisotrpoy, nonspatial random effects, analysis of variance, and more. Throughout, we use \pkg{spmodel} to fit SPGLMs to moose presence and counts in Alaska, United States (US), skewed conductivity data in the Southwestern US, harbor seal abundance trends in Alaska, US, and voter turnout rates in Texas, US.
keywords:
  # at least one keyword must be supplied
  formatted: [autoregressive model, geostatistical model, spatial covariance, spatial correlation]
  plain:     [autoregressive model, geostatistical model, spatial covariance, spatial correlation]
preamble: |
  \usepackage{amsmath,amsfonts,amssymb}
  \usepackage{bm, bbm}
  \usepackage{lineno}
  \usepackage{caption, subcaption}
  \usepackage{autonum}
output: rticles::jss_article
editor_options: 
  chunk_output_type: console
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ', width = 70)
library(spmodel)
library(ggplot2)
library(dplyr)
#library(emmeans)
#library(car)
library(here)
fig_path <- here("inst", "manuscript", "figures")
```

\linenumbers

\newpage

# Introduction {#sec-intro}

In practice, non-Gaussian data (e.g., binary, count, skewed, and proportion data) are ubiquitous. Non-Gaussian that belong to an exponential family data can be naturally modeled using a generalized linear model (GLM) regression framework [@nelder1972generalized; @mccullagh1989generalized]. In a GLM, an $n \times 1$ response variable $\mathbf{y}$ belongs to a statistical distribution (e.g., Binomial, Poisson) with some mean and variance. Often, the analysis goal is to study the impact of a linear function of several explanatory variables on $\mathbf{y}$ through a GLM. In this context, the latent (i.e., unobserved) mean of $\mathbf{y}$, $\boldsymbol{\mu}$, is linked to these explanatory variables via a link function:
\begin{align}\label{eq-glm}
f(\boldsymbol{\mu}|\mathbf{X}, \boldsymbol{\beta}) \equiv \mathbf{w} = \mathbf{X} \boldsymbol{\beta},
\end{align}
where for a sample size $n$, $f(\cdot)$ is a link function that connects $\boldsymbol{\mu}$ to $\mathbf{w}$, $\mathbf{X}$ is the $n \times p$ design matrix of explanatory variables, and $\boldsymbol{\beta}$ is the $p \times 1$ vector of fixed effects. While the mean is typically constrained in some way (e.g., if a probability, between zero and one), the link function generally makes $\mathbf{w}$ unconstrained. Common link functions include the log odds (i.e., logit) link for binary and proportion data and the log link for count and skewed data. Equation$~$\ref{eq-glm} can also be written in terms of the inverse link function, $f^{-1}(\cdot)$:
\begin{align}\label{eq-glm2}
\boldsymbol{\mu}|\mathbf{X}, \boldsymbol{\beta} \equiv f^{-1}(\mathbf{w}) = f^{-1}(\mathbf{X} \boldsymbol{\beta}).
\end{align}

The GLM fixed effects ($\boldsymbol{\beta}$) are typically estimated via maximum likelihood [@chambers1992S]. It is often convenient to compute the maximum likelihood estimates using the iteratively reweighted least squares (IRWLS) algorithm [@wood2017generalized], which is the approach used by the  \code{glm()} function in the \proglang{R} programming language [@rcore2024]. GLMs add an additional layer of complexity compared to linear regression models, as the left-hand size of Equation$~$\ref{eq-glm} is a function of the mean of $\mathbf{y}$ rather than $\mathbf{y}$ itself (as in linear regression models).

The standard GLM assumes the elements of $\mathbf{y}$ are independent. This independence assumption is typically impractical for spatial data. For spatial data, nearby observations tend to be more similar than distant observations [@tobler1970computer], which leads to positive spatial covariance. The consequences of ignoring spatial covariance in statistical models for spatial data can be severe and include imprecise parameter estimates as well as misleading standard errors that inflate Type-I error rates and decrease power [@zimmerman2024spatial].

An approach for handling spatial data using a GLM is to assume the elements of $\mathbf{w}$ share covariance that varies spatially. This is achieved by adding to Equation$~$\ref{eq-glm} two random effects, $\boldsymbol{\tau}$ and $\boldsymbol{\epsilon}$. The random effect $\boldsymbol{\tau}$ is an $n \times 1$ column vector of spatially dependent random errors. We assume that $\text{E}(\boldsymbol{\tau}) = \boldsymbol{0}$ and $\text{Cov}(\boldsymbol{\tau}) = \sigma^2_\tau \mathbf{R}$, where $\text{E}(\cdot)$ and $\text{Cov}(\cdot)$ denote expectation and covariance, respectively. The variance parameter $\sigma^2_\tau$ controls the magnitude of spatial covariance and is often called a partial sill. The matrix $\mathbf{R}$ is an $n \times n$ spatial correlation matrix that depends on a range parameter controlling the distance-decay rate of the spatial correlation. One example of a spatial covariance matrix is the "exponential", which is given by
\begin{align}\label{eq-spcov-exp}
  \text{Cov}(\boldsymbol{\tau}) = \sigma^2_{de} \mathbf{R}_{exp} =  \sigma^2_{de} \exp(-\mathbf{H}/\phi),
\end{align}
where $\mathbf{H}$ is a matrix of pairwise distances among the elements of $\mathbf{y}$ and $\phi$ is the range parameter. From Equation$~$\ref{eq-spcov-exp}, as the distance between two elements of $\mathbf{y}$ increases, the spatial covariance decreases, which reflects intuition. Moreover, as the range parameter, $\phi$, increases, the strength of spatial dependence increases (Figure$~$\ref{fig-range}). The random effect $\boldsymbol{\epsilon}$ is an $n \times 1$ column vector of independent random errors. We assume that $\text{E}(\boldsymbol{\epsilon}) = \boldsymbol{0}$ and $\text{Cov}(\boldsymbol{\tau}) = \sigma^2_\epsilon \mathbf{I}$, where $\mathbf{I}$ is an $n \times n$ identity matrix. The variance parameter $\sigma^2_\epsilon$ controls the magnitude of nonspatial variability (i.e., fine-scale variation) and is often called a nugget.  Often in spatial statistics, quantities are explicitly referenced with respect to $\mathbf{s}$, a vector of coordinates indexing the observation [@cressie1993statistics]. For example, $\mathbf{y}$ and $\mathbf{X}$ may instead be written $\mathbf{y}(\mathbf{s})$ and $\mathbf{X}$, respectively. We acknowledge the utility of this nomenclature but drop the explicit dependence on $\mathbf{s}$ for simplicity of notation.

\begin{figure}
\centering
\includegraphics[width = 0.7\linewidth]{figures/figure-01.png}
\caption{An exponential spatial correlation function with varying range parameters.}
\label{fig-range}
\end{figure}

Through inclusion of $\boldsymbol{\tau}$ and $\boldsymbol{\epsilon}$, the spatial GLM (SPGLM) can be written as
\begin{align}\label{eq-spglm}
f(\boldsymbol{\mu}|\mathbf{X}, \boldsymbol{\beta}, \boldsymbol{\tau}, \boldsymbol{\epsilon}) \equiv \mathbf{w} = \mathbf{X} \boldsymbol{\beta} + \boldsymbol{\tau} + \boldsymbol{\epsilon}.
\end{align}
Assuming independence among $\boldsymbol{\tau}$ and $\boldsymbol{\epsilon}$, it follows that 
\begin{align}\label{eq-spcov}
 \text{Cov}(\boldsymbol{\tau} + \boldsymbol{\epsilon}) = \text{Cov}(\boldsymbol{\tau}) + \text{Cov}(\boldsymbol{\epsilon}) = \sigma^2_{\tau}\mathbf{R} + \sigma^2_{\epsilon} \mathbf{I}.
\end{align}
Henceforth, we refer to $\sigma^2_{\tau}$ as $\sigma^2_{de}$ (for spatially dependent error variance) and $\sigma^2_{\epsilon}$ as $\sigma^2_{ie}$ (for independent error variance). The parameters $\sigma^2_{de}$, $\sigma^2_{ie}$, and $\phi$, in addition to any other parameters in $\mathbf{R}$, compose $\boldsymbol{\theta}$, the covariance parameter vector.

Fitting and using SPGLMs is challenging both conceptually and computationally [@bolker2009generalized]. Recently, however, there have been numerous, significant advances in \proglang{R} software that have made these models more accessible to practitioners. The \pkg{brms} [@burkner2017brms], \pkg{carBayes} [@lee2013carbayes], \pkg{ngspatial} [@hughes2020ngspatial], \pkg{R-INLA} [@lindgren2015bayesian] and \pkg{inlabru} [@bachl2019inlabru], \pkg{spBayes} [@finley2007spbayes], \pkg{spOccupancy} [@doser2022spoccupancy], \pkg{spAbundance} [@doser2024spabundance], and \pkg{spNNGP} [@finley2002spnngp] packages take a Bayesian approach, either directly sampling from posterior distributions of parameters (e.g., using MCMC) or approximating them. A benefit of Bayesian approaches is that prior information can be incorporated and uncertainty quantification of parameter estimates is straightforward. However, Bayesian approaches, especially those using MCMC, can be computationally expensive. In order to reduce computation time, many of these packages work with the precision matrix instead of the covariance matrix so that computationally expensive matrix inversion is not required. For example, \pkg{R-INLA} uses the precision matrix and tends to be very fast. Working with precision matrices, however, can be more restrictive and less intuitive than working directly with the covariance matrix. The \pkg{FRK} [@sainsbury2024modeling], \pkg{glmmTMB} [@brooks2017glmmtmb], \pkg{hglm} [@ronnegard2010hglm], \pkg{mgcv} [@wood2017generalized], and \pkg{spaMM} [@rousset2014spamm] packages directly use Laplace, quasi-likelihood, or reduced-rank approaches to estimate parameters. These direct approaches tend to be computationally efficient, as they don't rely on MCMC sampling. In contrast to the Bayesian approach, a drawback of these direct approaches is that prior information cannot be formally incorporated and covariance parameter uncertainty is more challenging to quantify. The \pkg{sdmTMB} [@anderson2024sdmtmb] package combines elements of \pkg{R-INLA}, \pkg{glmmTMB}, and Gaussian Markov random fields to fit a wide variety of SPGLMs, while \pkg{tinyVAST} [@thorson2025tinyVAST] extends some of these models to multivariate or (dynamic) structural equation models.

Building from @evangelou2011estimation and @bonat2016practical, @ver2024marginal proposed a novel approach for fitting SPGLMs that leverages the Laplace approximation while marginalizing over both the latent $\mathbf{w}$ and the fixed effects ($\boldsymbol{\beta}$) and accommodating spatial covariance. This approach performed efficiently in a variety of simulation settings, generally having appropriate confidence interval coverage for the fixed effects and prediction interval coverage for $\mathbf{w}$ at new locations. The approach performed similarly to the Bayesian SPGLM approach in \pkg{spBayes} and the automatic differentiation SPGLM approach in \pkg{glmmTMB} but was much faster. At small sample sizes, the approach outperformed the approximate Bayesian SPGLM approach in \pkg{R-INLA} and had similar computational times. For moderate sample sizes, it performed similarly to \pkg{R-INLA}, though \pkg{R-INLA} was faster. The novel Laplace approach is particularly attractive for two reasons. First, it is general enough that can be applied to any covariance structure (not just spatial). Second, after estimating the covariance parameters, analytical solutions exist for the fixed effects (and their standard errors) as well as predictions of the latent $\mathbf{w}$ at new locations (and their standard errors).

The \pkg{spmodel} \proglang{R} package [@dumelle2023spmodel] recently provided a full set of modeling tools for SPGLMs fit using the novel Laplace approach described by @ver2024marginal. These modeling tools are approachable and mirror the familiar \code{glm()} syntax from base-\proglang{R}, making the transition from GLMs to SPGLMs relatively seamless. The \code{spglm()} function fits SPGLMs for point-referenced data (e.g., x-coordinates and y-coordinates representing point locations in a field; models are sometimes called "geostatistical" models), while the \code{spgautor()} function fits SPGLMs for areal data (e.g., polygon boundaries representing geographic subsets of a region; models are sometimes called "autoregressive" models). \pkg{spmodel} supports the binomial distribution for binary data, Poisson and negative binomial distributions for count data, Gamma and inverse Gaussian distributions for skewed data, and the beta distribution for proportion data. There are 20 different spatial covariance structures available including the exponential, Gaussian, and spherical for point-referenced data (Figure$~$\ref{fig-type}) and the conditional autoregressive, and simultaneous autoregressive structures for areal data. \pkg{spmodel} provides tools for commonly used model summaries, visualizations, and diagnostics (e.g., Cook's distance) using standard \proglang{R} helper functions like \code{summary()}, \code{plot()}, \code{fitted()}, and \code{tidy()}, among othres. \pkg{spmodel} also provides tools to predict $\mathbf{w}$ at new locations and quantify uncertainty in those prediction using \code{predict()} and \code{augment()}. This core functionality, combined with several advanced features we describe throughout the manuscript, enables \pkg{spmodel} to introduce novel, important SPGLM modeling tools previously missing from the existing \proglang{R} ecosystem.

\begin{figure}
\centering
\includegraphics[width = 0.7\linewidth]{figures/figure-02.png}
\caption{Exponential, Gaussian, and spherical spatial correlation functions all with range parameters equal to 0.5.}
\label{fig-type}
\end{figure}

\pkg{spmodel} (version 0.11.0) is arguably most similar to \pkg{sdmTMB} (version 0.7.4) in terms of scope and feel. Both packages use similar syntax as \code{glm()}, accommodate flexible \code{formula} arguments (e.g., offsets, splines), handle spatial covariance that decays at different rates in different directions (i.e., geometric anisotropy), incorporate nonspatial random effects, support other \proglang{R} packages for modeling like \pkg{broom} [@robinson2021broom; @kuhn2022tidy], \pkg{emmeans} [@lenth2024emmeans], and \pkg{car} [@fox2019car], and have tools for model summaries, prediction, and simulating data. There are some notable differences between the two packages, however. \pkg{sdmTMB} supports several additional GLM distributions like the Tweedie, supports Hurdle models, and can incorporate prior information through Bayesian applications. \pkg{sdmTMB} also provides tools for working with temporal data and spatiotemporal data and provides enhanced visualizations of the model's marginal effects. \pkg{sdmTMB} does require a preprocessing step of constructing a mesh for the stochastic partial differential equation approach, and the density of the mesh can affect model results and computational complexity. On the other hand, \pkg{spmodel} does not require the construction of a mesh prior to modeling. \pkg{spmodel} also supports 20 different spatial covariances and models them directly, rather than using a precision matrix approximation to the Matérn spatial covariance as in \pkg{sdmTMB}. \pkg{spmodel} can also model areal data directly using neighborhood distance and autoregressive models rather than relying on the polygon centroid (as in \pkg{sdmTMB}), which may not be within the polygon's boundaries. \pkg{spmodel} also provides experimental design tools (e.g., analysis of variance, contrasts), supports \pkg{sf} objects in modeling and prediction functions [@pebesma2018sf], has several specialized model diagnostics like leverage values and Cook's distances, and has analytic solutions for fixed effect and prediction standard errors. Other similarities and differences do exist between \pkg{sdmTMB} and \pkg{spmodel}, and both packages continue to evolve. Overall, we believe that these packages are complementary and enhance the suite of SPGLM tools accessible to practitioners.

The rest of this article is organized as follows. In Section$~$\ref{sec-spglm}, we provide some background for the SPGLM fitting and prediction routines in \pkg{spmodel}. In Section$~$\ref{sec-applications}, we provide an overview of core SPGLM functionaility in \pkg{spmodel} by modeling moose presence in Alaska, United States (US). In Section$~$\ref{sec-applications2}, we model moose counts in Alaska, US; skewed lake conductivity in the Southwestern US; harbor seal abundance trend behavior in Alaska, US; and voter turnout rates in Texas, US. And in Section$~$\ref{sec-discussion}, we end with a discussion synthesizing \pkg{spmodel}'s contributions to the analysis of SPGLMs in \proglang{R}.

# The spatial generalized linear model and marginalization {#sec-spglm}

The novel Laplace approach implemented in \pkg{spmodel} formally maximizes a hierarchical GLM likelihood [@lee1996hierarchical; @wood2017generalized], making likelihood-based statistics for model comparison like AIC [@akaike1974new], AICc [@hoeting2006model], BIC [@schwarz1978estimating], deviance [@mccullagh1989generalized], and likelihood ratio tests available. These types of statistics are not available for quasi-likelihood [@wedderburn1974quasi; @breslow1993approximate] or pseudo-likelihood approaches [@wolfinger1993generalized], which only specify the first two moments of a distribution. Next, we describe a brief overview of the approach and how it can be used for several primary data analysis tasks [@tredennick2021practical] like model comparison, parameter estimation, inference, model diagnostics, and prediction.

## Formulating the hierarchical likelihood

We can write the SPGLM likelihood hierarchically as
\begin{align}\label{eq-marginal}
  [\mathbf{y}|\mathbf{X}, \varphi, \boldsymbol{\theta}] = \int_{\mathbf{w}} \int_{\boldsymbol{\beta}} [\mathbf{y} | f^{-1}(\mathbf{w}), \varphi] [\mathbf{w} | \mathbf{X}, \boldsymbol{\theta}] d\boldsymbol{\beta} d\mathbf{w},
\end{align}
where $[\mathbf{y} | f^{-1}(\mathbf{w}), \varphi]$ is the density for the appropriate response distribution of $\mathbf{y}$ (e.g., binomial, Poisson) given the latent $\mathbf{w}$ and dispersion parameter ($\varphi$), and $[\mathbf{w} | \mathbf{X}, \boldsymbol{\theta}]$ is the multivariate Gaussian density for $\mathbf{w}$ given the explanatory variables ($\mathbf{X}$), fixed effects ($\boldsymbol{\beta}$), and spatial covariance parameters ($\boldsymbol{\theta}$). The elements of $[\mathbf{y} | f^{-1}(\mathbf{w}), \varphi]$ are conditionally independent (given $\mathbf{w}$), but the elements of $[\mathbf{w} | \mathbf{X}, \boldsymbol{\theta}]$ share spatial covariance. Following @harville1977maximum, we can integrate $\boldsymbol{\beta}$ out of Equation$~$\ref{eq-marginal}, which yields
\begin{align}\label{eq-marginal2}
  [\mathbf{y}|\mathbf{X}, \varphi, \boldsymbol{\theta}] = \int_{\mathbf{w}} [\mathbf{y} | f^{-1}(\mathbf{w}), \varphi] [\mathbf{w} | \mathbf{X}, \boldsymbol{\theta}] d\mathbf{w},
\end{align}
where $[\mathbf{w} | \mathbf{X}, \boldsymbol{\theta}]$ is the restricted (i.e., residual) multivariate Gaussian density [@patterson1971recovery] for $\mathbf{w}$ given the explanatory variables and covariance parameters. Equation$~$\ref{eq-marginal2} can synonymously be written after profiling the overall variance out of $\boldsymbol{\Sigma}$, which reduces the dimension of $\boldsymbol{\theta}$ by one for optimization [@wolfinger1994computing]. The restricted multivariate Gaussian density is given by
\begin{align}\label{eq-reml-def}
[\mathbf{w} | \mathbf{X}, \boldsymbol{\theta}] = \frac{\exp(-\frac{1}{2}(\mathbf{y} - \mathbf{X}\tilde{\boldsymbol{\beta}}) \boldsymbol{\Sigma}^{-1} (\mathbf{y} - \mathbf{X}\tilde{\boldsymbol{\beta}})^T)}{(2 \pi)^{(n - p)/2} |\boldsymbol{\Sigma}|^{1/2}|\mathbf{X}^T \boldsymbol{\Sigma}^{-1} \mathbf{X}|^{1/2}},
\end{align}
where $\tilde{\boldsymbol{\beta}} = (\mathbf{X}^T \boldsymbol{\Sigma}^{-1} \mathbf{X})^{-1} \mathbf{X}^T \boldsymbol{\Sigma}^{-1} \mathbf{w}$ and $|\cdot|$ denotes the determinant. Next, let 
\begin{align}\label{eq-marginal03}
  \ell_\mathbf{w} = \log([\mathbf{y} | f^{-1}(\mathbf{w}), \varphi] [\mathbf{w} | \mathbf{X}, \boldsymbol{\theta}])
\end{align}
and rewrite Equation$~$\ref{eq-marginal2} as 
\begin{align}\label{eq-marginal3}
  [\mathbf{y}|\mathbf{X}, \varphi, \boldsymbol{\theta}] = \int_{\mathbf{w}} \exp(\ell_\mathbf{w}) d\mathbf{w}.
\end{align}
A second-order Taylor series expansion of $\ell_\mathbf{w}$ around a point $\hat{\mathbf{w}}$ yields
\begin{align}\label{eq-marginal4}
  [\mathbf{y}|\mathbf{X}, \varphi, \boldsymbol{\theta}] \approx \int_{\mathbf{w}} \exp(\ell_{\hat{\mathbf{w}}} + \mathbf{g}^T(\mathbf{w} - \hat{\mathbf{w}}) + \frac{1}{2}(\mathbf{w} - \hat{\mathbf{w}})^T \mathbf{G} (\mathbf{w} - \hat{\mathbf{w}}))d\mathbf{w},
\end{align}
where $\mathbf{g}$ and $\mathbf{G}$ are the gradient and Hessian, respectively, of $\ell_\mathbf{w}$ with respect to $\mathbf{w}$. If $\hat{\mathbf{w}}$ is a value for which $\mathbf{g} = \mathbf{0}$,
\begin{align}\label{eq-marginal5}
  [\mathbf{y}|\mathbf{X}, \varphi, \boldsymbol{\theta}] \approx \exp(\ell_{\hat{\mathbf{w}}}) \int_{\mathbf{w}} \exp(-\frac{1}{2}(\mathbf{w} - \hat{\mathbf{w}})^T (-\mathbf{G}) (\mathbf{w} - \hat{\mathbf{w}}))d\mathbf{w}.
\end{align}
The integral in Equation$~$\ref{eq-marginal5} can be solved by leveraging properties of the normalizing constant of a multivariate Gaussian distribution. Thus, rewriting $\exp(\ell_{\hat{\mathbf{w}}})$ yields
\begin{align}\label{eq-marginal6}
  [\mathbf{y}|\mathbf{X}, \varphi, \boldsymbol{\theta}] \approx [\mathbf{y} | f^{-1}(\hat{\mathbf{w}}), \varphi] [\hat{\mathbf{w}} | \mathbf{X}, \boldsymbol{\theta}] (2 \pi)^{n/2}|-\mathbf{G}_{\hat{\mathbf{w}}}|^{-1/2}.
\end{align}

Maximizing the natural logarithm of Equation$~$\ref{eq-marginal6} requires a doubly iterative process over $\boldsymbol{\theta}$ and $\varphi$ as well as $\mathbf{w}$, eventually yielding the the marginal restricted maximum likelihood estimators $\hat{\varphi}$ and $\hat{\boldsymbol{\theta}}$ and their corresponding values of $\hat{\mathbf{w}}$. Maximizing this log likelihood is a computationally expensive operation that involves repeatedly evaluating $\boldsymbol{\Sigma}^{-1}$, $\mathbf{g}$, and $\mathbf{G}$; see @ver2024marginal for more details and forms of $\mathbf{g}$ and $\mathbf{G}$ for various response distributions.

## Estimating fixed effects

Though the fixed effects are integrated out of the likelihood, we can still estimate them using generalized least squares (GLS) principles, a common practice for linear models estimated using restricted maximum likelihood methods. Had we observed $\mathbf{w}$, a GLS estimator for $\boldsymbol{\beta}$ is given by
\begin{align}\label{eq-gls1}
  \hat{\boldsymbol{\beta}} = (\mathbf{X}^\top \boldsymbol{\Sigma}^{-1}\mathbf{X})^{-1}\mathbf{X}^\top \boldsymbol{\Sigma}^{-1}\mathbf{w} = \mathbf{B}\mathbf{w},
\end{align}
where $\mathbf{B} = (\mathbf{X}^\top \boldsymbol{\Sigma}^{-1}\mathbf{X})^{-1}\mathbf{X}^\top \boldsymbol{\Sigma}^{-1}$. However, we only observe $\hat{\mathbf{w}}$, so it is reasonable to define $\hat{\boldsymbol{\beta}} = \mathbf{B}\hat{\mathbf{w}}$. Thus, to derive properties of $\hat{\boldsymbol{\beta}}$ like expectation and variance, we must derive these properties for $\hat{\mathbf{w}}$. To do so, we must condition on $\mathbf{w}$ as if it were observed and invoke properties of the laws of total expectation and variance. Because $\hat{\mathbf{w}}$ was optimized via the likelihood, we assume that given $\mathbf{w}$, $\hat{\mathbf{w}}$ has mean $\mathbf{w}$ and variance approximately equal to $-\mathbf{H}^{-1}$ (the inverse Hessian). It follows that $\text{E}(\hat{\mathbf{w}})$ is given by
\begin{align}
  \text{E}(\hat{\mathbf{w}}) = \text{E}(\text{E}(\hat{\mathbf{w}} | \mathbf{w}))
   = \text{E}(\mathbf{w}) 
   = \mathbf{X}\boldsymbol{\beta}
\end{align}
and $\text{Var}(\hat{\mathbf{w}})$ is given by
\begin{align}
  \text{Var}(\hat{\mathbf{w}}) & = \text{E}(\text{Var}(\hat{\mathbf{w}} | \mathbf{w})) + \text{Var}(\text{E}(\hat{\mathbf{w}} | \mathbf{w})) \\
  & = \text{E}(-\mathbf{H}^{-1}) + \text{Var}(\mathbf{w})\\
  & = -\mathbf{H}^{-1} + \boldsymbol{\Sigma}
\end{align}
Putting this all together, it follows that
\begin{align}
  \text{E}(\hat{\boldsymbol{\beta}})  = \text{E}(\mathbf{B}\hat{\mathbf{w}}) 
  = \mathbf{B}\text{E}(\hat{\mathbf{w}}) =  (\mathbf{X}^\top \boldsymbol{\Sigma}^{-1}\mathbf{X})^{-1}(\mathbf{X}^\top \boldsymbol{\Sigma}^{-1} \mathbf{X})\boldsymbol{\beta} = \boldsymbol{\beta}
\end{align}
and
\begin{align}
  \text{Var}(\hat{\boldsymbol{\beta}}) & = \text{Var}(\mathbf{B}\hat{\mathbf{w}}) \\
  & = \mathbf{B} \text{Var}(\hat{\mathbf{w}}) \mathbf{B}^\top\\
  & = \mathbf{B} (-\mathbf{H}^{-1} + \boldsymbol{\Sigma}) \mathbf{B}^\top \\
  & = \mathbf{B}(-\mathbf{H})^{-1}\mathbf{B}^\top + \mathbf{B}\boldsymbol{\Sigma}\mathbf{B}^\top \\
  & = \mathbf{B}(-\mathbf{H})^{-1}\mathbf{B}^\top + (\mathbf{X}^\top \boldsymbol{\Sigma}^{-1}\mathbf{X})^{-1}
\end{align}
In practice, $\text{Var}(\hat{\boldsymbol{\beta}})$ is estimated by evaluating $\boldsymbol{\Sigma}$ at $\hat{\boldsymbol{\theta}}$, the estimated covariance parameter vector.

These results are important because they justify closed-form solutions for $\hat{\boldsymbol{\beta}}$ and its associated variance. Closed-form solutions are useful because they bypass the need for sampling-based strategies to evaluate the mean and variance of $\hat{\boldsymbol{\beta}}$, a common technique for other approaches to SPGLMs like Bayesian MCMC.

## Inspecting model diagnostics

Inspecting model diagnostics is an important step of the modeling process that can yield valuable insights into model behavior and unusual observations. @montgomery2021introduction contextualize three components of unusual observations: outliers, leverage, and influence. An observation is an outlier if it has an unusual response value relative to expectation. The response GLM residuals simply compare the observation to its fitted latent mean:
\begin{align}
  \mathbf{r}_{r} = \mathbf{y} - f^{-1}(\hat{\mathbf{w}})
\end{align}
Because observations often have a unique support in a GLM (e.g., only two possible response values for binary data) and the variance of an observation generally depends on its mean, response residuals lack some utility. Deviance residuals are a function of response residuals that are appropriately scaled to behave more like response residuals in a standard linear model. Deviance residuals are given by
\begin{align}
  \mathbf{r}_{d} = sign(\mathbf{r}_{r})\sqrt{\mathbf{d}},
\end{align}
where $\mathbf{d}$ is a vector of individual deviances. The sum of the squared deviance residuals equals the sum of $\mathbf{d}$. The sum of $\mathbf{d}$ is the deviance of the model fit, which quantifies twice the difference in log likelihoods between the a saturated model that fits every observation perfectly (i.e., $\mathbf{y} = f^{-1}(\hat{\mathbf{w}}_i)$ for all $i$) and the fitted model [@myers2012generalized]. Deviance is often used as a fit statistic; lower values of deviance imply a better model fit. Pearson and standardized residuals are other types of GLM residuals that involve a scaling of the response residuals; the Pearson residuals scale $\mathbf{r}_{r}$ by the square root of $\mathbf{V}$, while the standardized residuals scale the deviance residuals by $\frac{1}{\sqrt{(1 - \mathbf{L}_{ii})}}$, where $\mathbf{L}_{ii}$ is the $i$th diagonal element of the leverage matrix, which we discuss next.

An observation has high leverage if its combination of explanatory variables is far away from other observations. In a linear model, the leverage (i.e., hat) values are the diagonal of the leverage (i.e., projection, hat) matrix, $\mathbf{L} = \mathbf{X}(\mathbf{X}^\top \mathbf{X})^{-1}\mathbf{X}^\top$. In a GLM, the leverage matrix is given by
\begin{align}
  \mathbf{L} = \mathbf{V}^{1/2} \mathbf{X} (\mathbf{X}^\top \mathbf{V} \mathbf{X}) \mathbf{X}^\top \mathbf{V}^{1/2},
\end{align}
where $\mathbf{V}$ is a diagonal matrix with $i$th diagonal element equal to the variance of the response distribution evaluated at $f^{-1}(\mathbf{w}_i)$ [@faraway2016extending]; $\mathbf{V}$ is sometimes called the GLM weight matrix. The larger the value of $\mathbf{L}_{ii}$, the more severe the leverage from the $i$th observation.

An observation is influential if it has a sizable impact on model fit. Influence is measured using Cook's distance [@cook1979influential; @cook1982residuals], which is given for a GLM by
\begin{align}
  \mathbf{c} = \frac{\mathbf{r}^2_{s}}{\text{tr}(\mathbf{L})} \frac{diag(\mathbf{L})}{(\mathbf{1} - diag(\mathbf{L}))},
\end{align}
where $\mathbf{r}^2_{s}$ are the standardized residuals and $diag(\mathbf{L})$ indicates the diagonal elements of the leverage matrix. The larger the value of $\mathbf{c}_i$, the more severe the influence from the $i$th observation.  @montgomery2021introduction provide guidance for interpreting these types of statistics, including cutoffs to consider when identifying unusual residual, leverage, or influence values.

In a linear model, the $R^2$ (R-squared) statistic quantifies the proportion of variability in the data captured by the explanatory variables. It is calculated as one minus the ratio of the error sum of squares to the total sum of squares [@rencher2008linear]. In a GLM, there are many ways to define a statistic that emulates the aforementioned meaning of $R^2$ from the linear model [@smith2013comparison]. This statistic is called a pseudo R-squared ($PR^2$). One $PR^2$ for GLMs simply replaces the sums of squares ratio from the linear model with the deviance ratio:
\begin{align}
  PR^2 = 1 - \frac{deviance_{fit}}{deviance_{null}},
\end{align}
where $deviance_{fit}$ is the deviance of the fitted model (sometimes called the residual deviance) and $deviance_{null}$ is the deviance of the model taking $\mathbf{X} \equiv \mathbf{1}$, a column of all ones (i.e., an intercept-only model).  In practice, $deviance_{null}$ is derived by computing $\hat{\mathbf{w}}$ when $\mathbf{X} \equiv \mathbf{1}$ given $\hat{\boldsymbol{\theta}}$ and $\hat{\varphi}$ from the fitted model. Like $R^2$, $PR^2$ can be adjusted to account for the numbers of parameters estimated in a model. Because the $deviance_{null}$ denominator changes across fitted models (as the values of $\hat{\boldsymbol{\theta}}$ and $\hat{\varphi}$ change), this statistic should not be used as a model comparison tool. Rather, it should be used as an informative diagnostic tool that unique to each model fit and describes how much variability from that model is attributable to the explanatory variables.

## Predicting at new locations

We may also predict values of the latent mean (on the link scale) at new locations by leveraging the spatial covariance between observed locations and new locations (spatial prediction is also called Kriging; see @cressie1990origins). Again suppose that we observed $\mathbf{w}$ and we want to make predictions at $\mathbf{u}$, a vector of latent means at the new locations that follows the same SPGLM from Equation$~$\ref{eq-spglm} and having fixed effects design matrix, $\mathbf{X}_{\mathbf{u}}$. The vector $(\mathbf{w}, \mathbf{u})^\top$ has expectation $(\mathbf{X}\boldsymbol{\beta}, \mathbf{X}_\mathbf{u}\boldsymbol{\beta})^\top$ and covariance matrix $\begin{bmatrix} \boldsymbol{\Sigma} & \boldsymbol{\Sigma}_{\mathbf{w}\mathbf{u}} \\ \boldsymbol{\Sigma}_{\mathbf{u}\mathbf{w}} & \boldsymbol{\Sigma}_{\mathbf{u}\mathbf{u}} \end{bmatrix}$, where $\boldsymbol{\Sigma} = \text{Var}(\mathbf{w}, \mathbf{w})$, $\boldsymbol{\Sigma}_{\mathbf{w} \mathbf{u}} = \text{Var}(\mathbf{w}, \mathbf{u})$, $\boldsymbol{\Sigma}_{\mathbf{u} \mathbf{w}} = \boldsymbol{\Sigma}_{\mathbf{w} \mathbf{u}}^\top$ and $\boldsymbol{\Sigma}_{\mathbf{u}, \mathbf{u}} = \text{Var}(\mathbf{u}, \mathbf{u})$
By assumption, we have observed $\mathbf{w}$, so we may derive the conditional distribution of $\mathbf{u}|\mathbf{w}$, which has the following properties:
\begin{align}
  \text{E}(\mathbf{u} | \mathbf{w}) & = \mathbf{X}_{\mathbf{u}} \boldsymbol{\beta} + \boldsymbol{\Sigma}_{\mathbf{u}, \mathbf{w}}\boldsymbol{\Sigma}^{-1}(\mathbf{w} - \mathbf{X}\boldsymbol{\beta}) \\
  \text{E}(\mathbf{u} | \mathbf{w}) & = \boldsymbol{\Sigma}_{\mathbf{u}, \mathbf{u}} - \boldsymbol{\Sigma}_{\mathbf{u}, \mathbf{w}} \boldsymbol{\Sigma}^{-1} \boldsymbol{\Sigma}_{\mathbf{w}, \mathbf{u}}
\end{align}
@ver2024marginal show how these equations are adjusted to reflect uncertainty in both $\hat{\boldsymbol{\beta}}$ and $\hat{\mathbf{w}}$ while leveraging the laws of total expectation and variance yet again. They derive the predictor of $\mathbf{u}$, $\hat{\mathbf{u}}$, and its associated variance, given by:
\begin{align}
  \hat{\mathbf{u}} & = \mathbf{X}_{\mathbf{u}} \hat{\boldsymbol{\beta}} + \boldsymbol{\Sigma}_{\mathbf{u}, \mathbf{w}}\boldsymbol{\Sigma}^{-1}(\hat{\mathbf{w}} - \mathbf{X}\hat{\boldsymbol{\beta}}) \\
  \text{Var}(\hat{\mathbf{u}}) & = \boldsymbol{\Sigma}_{\mathbf{u}, \mathbf{u}} - \boldsymbol{\Sigma}_{\mathbf{u}, \mathbf{w}} \boldsymbol{\Sigma}^{-1} \boldsymbol{\Sigma}_{\mathbf{w}, \mathbf{u}} + \mathbf{K}(\mathbf{X}^\top \boldsymbol{\Sigma}^{-1}\mathbf{X})^{-1}\mathbf{K}^\top + \boldsymbol{\Lambda}(-\mathbf{H})^{-1}\boldsymbol{\Lambda}^\top,
\end{align}
where $\mathbf{K} = \mathbf{X}_{\mathbf{u}} - \boldsymbol{\Sigma}_{\mathbf{u}, \mathbf{w}} \boldsymbol{\Sigma}^{-1} \mathbf{X}$ and $\boldsymbol{\Lambda} = \mathbf{X}_{\mathbf{u}}\mathbf{B} + \boldsymbol{\Sigma}_{\mathbf{u}, \mathbf{w}}\boldsymbol{\Sigma}^{-1}(\mathbf{1} - \mathbf{X}\mathbf{B})$ for a vector of ones, $\mathbf{1}$. As with $\hat{\boldsymbol{\beta}}$, in practice these covariance matrices are evaluated at $\hat{\boldsymbol{\theta}}$.

# Modeling moose presence in Alaska, USA {#sec-applications}

The \code{moose} data in \pkg{spmodel} contain information on moose (Alces Alces) presence in the Togiak region of Alaska, USA.   \code{moose} is an \code{sf} object, a special data frame that is supplemented with spatial information using the \pkg{sf} package in \proglang{R} [@pebesma2018sf]. The first few rows of \code{moose} look like:

```{r}
head(moose)
```

There are five columns: \code{elev}, the numeric site elevation (meters); \code{strat} a stratification variable for sampling with two levels, \code{"L"} and \code{"M"}, which are categorized by landscape metrics at each site; \code{count}, the number of moose at each site; \code{presence}, a factor that indicates whether at least one moose was observed at each site (\code{0} implies no moose; \code{1} implies at least one moose); and \code{geometry}, the NAD83/Alaska Albers (EPSG: 3338) projected coordinate of each site (these data are point-referenced because each observation occurs at point coordinates and are represented by a \code{POINT} geometry. Moose are most prevalent in the southwestern and eastern parts of the Togiak region (Figure$~$\ref{fig-moose-data}).

\begin{figure}
\centering
\includegraphics[width = 0.70\linewidth]{figures/figure-1.png}
\caption{Moose presence in Alaska. Circles represent moose presence or absence (based on color) and triangles represent locations at which moose presence probability predictions are desired.}
\label{fig-moose-data}
\end{figure}

The \code{moose_preds} data in \pkg{spmodel} is an \code{sf} object with point locations at which moose presence predictions are desired. Like \code{moose}, \code{moose_preds} contains \code{elev} and \code{strat} for each site:
```{r}
head(moose_preds)
```

## Model Fitting

SPGLMs in \pkg{spmodel} are fit using the \code{spglm()} function. The \code{spglm()} function requires four arguments: \code{formula}, the relationship between the response and explanatory variables; \code{family}, the response distribution assumed for the repsonse variable; \code{data}, the data frame that contains the variables in \code{formula}, and \code{spcov_type}, the type of spatial covariance. The \code{formula}, \code{family}, and \code{data} arguments are the three required arguments to \code{glm()} for nonspatial GLMs. So, the transition from \code{glm()} to \code{spglm()}
simply requires one additional argument: \code{spcov_type}. When \code{data} is not an \code{sf} object, \code{spglm()} also requires the \code{xcoord} and \code{ycoord} arguments, which indicate the columns in \code{data} that represent projected the x- and y-coordinates, respectively.

We use \code{spglm()} to fit a spatial GLM (i.e., here, a spatial logistic regression) quantifying the effect of elevation and strata on moose presence:
```{r}
spbin <- spglm(
  formula = presence ~ elev + strat,
  family = binomial,
  data = moose,
  spcov_type = "exponential"
)
```

The \code{summary()} function returns a model summary that returns relevant information like the function call, deviance residuals, a coefficients table of fixed effects, the pseudo R-squared, spatial covariance parameter coefficient estimates, and the GLM dispersion parameter (fixed at one in logistic regression):
```{r}
summary(spbin)
```

The model suggests that elevation is positively associated with the log odds of moose presence ($p$-value $\approx$ 0.087), after controlling for strata. The model also suggests that moose have a higher log odds of presence in the \code{"M"} strata compared to the  \code{"L"} strata ($p$-value < 0.001), after controlling for elevation.

The fixed effects coefficients table from \code{summary()} is often of primary scientific interest, but it is not immediately usable when printed directly to the \proglang{R} console. The \code{tidy()} function tidies this table, turning it into a data frame (i.e., a tibble) with standard column names:
```{r}
tidy(spbin, conf.int = TRUE)
```

## Model Comparison

The strength of spatial covarinace in the data affects how beneficial a SPGLM is relative to a GLM. When the spatial covariance is strong, the SPGLM should notably outperform the GLM. When the spatial covariance is weak, the SPGLM and GLM should perform similarly. We can quantify the benefits of incorporating spatial covariance for a particular data set by comparing the fit of a SPGLM to a GLM. We can fit a GLM in \code{spmodel} by specifying \code{spcov_type = "none"}:

```{r}
bin <- spglm(
  formula = presence ~ elev + strat,
  family = binomial,
  data = moose,
  spcov_type = "none"
)
```

While the \code{spglm()} approach evaluates the HGLMM likelhood with $\sigma^2_{de} = 0$ and $\sigma^2_{ie} \approx 0$ instead of just the GLM likelihood, the parameter estimates and their standard errors are the same:

```{r}
bin_glm <- glm(
  formula = presence ~ elev + strat,
  family = binomial,
  data = moose,
)
round(coef(bin), digits = 4)
round(coef(bin_glm), digits = 4)
round(sqrt(diag(vcov(bin))), digits = 4)
round(sqrt(diag(vcov(bin_glm))), digits = 4)
```

However, using \code{spglm()} instead of \code{glm()} ensures that \pkg{spmodel} helper functions are available and that each of the \code{spglm()} models uses the same likelihood:
```{r}
glance(spbin)
glance(bin)
```

The likelihood-based statistics AIC, AICc, BIC, and deviance are much lower for the SPGLM, indicating a better fit relative to the GLM. We may also perform a likelihood ratio test (LRT) between the two models, as the GLM is a special case of the SPGLM (i.e., is nested within the SPGLM):
```{r}
anova(spbin, bin)
```

The LRT provides strong evidence that the SPGLM is preferred to the GLM ($p$-value < 0.001).

An alternative approach to model comparison is to use a cross-validation procedure [@james2013introduction]. The \code{loocv()} function performs leave-one-out cross validation, comparing the predicted mean (on the response scale) to the observed response variable for each hold-out observation, recomputing estimates of $\boldsymbol{\beta}$ each iteration Then, statistics like bias, mean-squared-prediction error (MSPE), and the square root of MSPE (RMSPE) can be used to evaluate models:

```{r}
loocv(spbin)
loocv(bin)
```

Both models have negligible bias, but the SPGLM has much lower MSPE and RMSPE than the GLM, indicating the SPGLM predictions are far more efficient. Three separate metrics (likelihood-based statistics, likelihood-ratio test, and leave-one-out cross validation) prefer the SPGLM to the GLM.

We can compare two SPGLMs with different spatial covariance functions using likelihood-based statistics and leave-one-out cross validation, but we can't use the LRT because generally, the spatial covariance functions aren't nested:
```{r}
spbin2 <- update(spbin, spcov_type = "gaussian")
glances(spbin, spbin2)
loocv(spbin)
loocv(spbin2)
```

The \code{"exponential"} spatial covariance (\code{spbin}) has a slightly lower (better) deviance but slightly higher (worse) AIC, AICc, and BIC than the \code{"gaussian"} spatial covariance (\code{spbin2}). Both spatial covariance functions have similar leave-one-out cross validation metrics, though the \code{"gaussian"} spatial covariance RMSPE is slightly lower (better). For practical purposes, these models fit similarly.

Often times in spatial statistics, the difference in model fit between the best spatial model and worst spatial model is much smaller than the difference in model fit between the worst spatial model and the nonspatial model, implying that accounting for some form of spatial covariance is very beneficial. Two spatial covariance functions to consider starting with are the exponential and Gaussian, which have quite different origin behaviors (Figure$~$\ref{fig-type}), something @stein1999interpolation argues is important to characterize accurately.

## Model Diagnostics

\code{spmodel} provides a suite of tools for model diagnostics. One is \code{augment()}, which augments the model data with diagnostics:
```{r}
augment(spbin)
```

The fitted values (\code{.fitted}) can be returned on either the link ($\hat{\mathbf{w}}$) or response ($f^{-1}(\hat{\mathbf{w}})$) scale and the residuals (\code{.resid}) can deviance, Pearson, or response residuals. The default fitted values are the link scale and the default residuals are deviance residuals. Also returned by \code{augment()} are the leverage (\code{.hat}), Cook's distance (\code{.cooksd}), and standardized residuals \code{.std.resid}. A benefit of using \code{augment()} when \code{data} is an \code{sf} object is that the output is also an \code{sf} object, which makes it straightforward to create spatial diagnostic plots (Figure$~$\ref{fig-sp-diagnostic}). Standard \proglang{R} helpers (e.g., \code{fitted()}, \code{residuals()}) are also available to extract model diagnostics from the model object.

\begin{figure}
\centering
\includegraphics[width = 1\linewidth]{figures/figure-2.png}
\caption{Moose presence model diagnostics, including leverage values (left) and standardized residuals (right).}
\label{fig-sp-diagnostic}
\end{figure}

The \code{plot()} function can also be used to return similar diagnostics as from \code{lm()} and \code{glm()} with additional tools for diagnosing spatial covariance. For example, we can inspect Cook's distance values and the empirical spatial covariance as a function of distance with (Figure$~$\ref{fig-sp-diagnostic2}):
```{r, eval = FALSE}
plot(spbin, which = c(4, 7))
```

\begin{figure}
\centering
\includegraphics[width = 1\linewidth]{figures/figure-3.png}
\caption{Moose presence model diagnostics, including Cook's distance and the fitted spatial covariance as a function of distance (right).}
\label{fig-sp-diagnostic2}
\end{figure}

The \code{varcomp()} function partitions model variability into several different components, helping to elucidate the model's structure:
```{r}
varcomp(spbin)
```

The pseudo R-squared ($PR^2$) is reported in the first row. The remaining variability ($1 - PR^2$) is allocated proportionally to \code{de} and \code{ie} according to $\sigma^2_{de}$ and $\sigma^2_{ie}$. This variability partitioning is a useful that helps quantify how much the explanatory variables, residual spatial variance, and residual nonspatial variance contribute to model fit, but as with $PR^2$, should not be used as a model comparison tool.

## Prediction

We can predict the probability of moose presence using \code{predict()}:
```{r}
predict(spbin, newdata = moose_preds)[1:5]
```

By default, predictions are returned on the link scale, but this can be changed to the response scale via \code{type}:
```{r}
predict(spbin, newdata = moose_preds, type = "response")[1:5]
```

Predictions on the response scale are visualized alongside the fitted values ($f^{-1}(\hat{\mathbf{w}})$) in Figure$~$\ref{fig-moose-fit}.

\begin{figure}
\centering
\includegraphics[width = 0.70\linewidth]{figures/figure-4.png}
\caption{Moose presence probability fitted values and predictions. Fitted values are represeneted by circles and predictions by triangles.}
\label{fig-moose-fit}
\end{figure}

Prediction intervals for the probability of moose presence (on the link scale) are returned by suppplying \code{interval}:
```{r}
predict(spbin, newdata = moose_preds, interval = "prediction")[1:5, ]
```

We can alternatively use \code{augment()} to augment the prediction data with predictions. Arguments to \code{predict()} can also be passed to \code{augment()}:
```{r}
augment(spbin, newdata = moose_preds, interval = "prediction")
```

By using \code{augment()} when \code{newdata} is an \code{sf} object, predictions and their corresponding uncertainties are readily available for spatial mapping (Figure$~$\ref{fig-moose-int}).

\begin{figure}
\centering
\includegraphics[width = 1\linewidth]{figures/figure-5.png}
\caption{Moose presence probability prediction intervals. 95\% prediction interval lower bound (left) and 95\% prediction interval upper bound (right).}
\label{fig-moose-int}
\end{figure}

# Additional applications {#sec-applications2}

Throughout the remainder of this section, we briefly highlight some additional \pkg{spmodel} capabilities for SPGLMs. In Section$~$\ref{sec-moose-count}, we fit Poisson and negative binomial models with and without geometric anisotropy for the point-referenced moose count data. In Section$~$\ref{sec-lake}, we fit a Gamma model to the point-referenced lake conductivity data, showing how to perform a spatial analysis of variance (ANOVA), fit contrasts for models with interactions and use maximum likelihood to compare two models with different explanatory variables. In Section$~$\ref{sec-seal}, we fit a binomial model to the areal harbor seal trend data with a nonspatial random effect. Finally in Section$~$\ref{sec-texas}, we fit beta models to Texas voter turnout data, which can be treated as point-referenced or areal.

## Modeling moose counts in Alaska, USA {#sec-moose-count}

In addition to moose presence, moose counts are also recorded in \code{moose} (Figure$~$\ref{fig-moose-data-count}.  The Poisson and negative binomial response distributions can be used to model SPGLMs for count data. The Poisson distribution mean is equal to its variance, while the negative binomial has an extra parameter to accommodate overdispersion (where the variance is larger than the mean). Using a spherical spatial covariance function, we may fit both a Poisson and negative binomial SPGLM changing the \code{family} argument:

```{r}
sppois <- spglm(
  formula = count ~ elev + strat,
  family = poisson,
  data = moose,
  spcov_type = "spherical"
)
spnb <- update(sppois, family = nbinomial)
```

\begin{figure}
\centering
\includegraphics[width = 0.70\linewidth]{figures/figure-6.png}
\caption{Moose counts in Alaska. Circles represent moose counts (based on color) and triangles represent locations at which mean count predictions are desired.}
\label{fig-moose-data-count}
\end{figure}

Because the Poisson and negative binomial distributions have the same support (nonnegative integers), we can compare them using  AIC, AICc, or BIC:
```{r}
BIC(sppois, spnb)
```

Implicit in our spatial covariance functions thus far has been an assumption of geometric isotropy. A spatial covariance function is geometrically isotropic if it decays with distance at the same rate in all directions (Figure$~$\ref{fig-tropy}; left). A spatial covariance is geometrically isotropic if it decays with distance at different rates in different directions (Figure$~$\ref{fig-tropy}; right). Geometric anisotropy is formally incorporated by rotating and scaling original coordinates, yielding transformed coordinates that are geometrically isotropic:
\begin{align}
  \begin{bmatrix}
    x^* \\
    y^*
  \end{bmatrix} = 
  \begin{bmatrix}
    1 & 0 \\
    0 & 1 / \omega
  \end{bmatrix}
  \begin{bmatrix}
    \cos(\alpha) & \sin(\alpha) \\
    -\sin(\alpha) & \cos(\alpha)
  \end{bmatrix}  
  \begin{bmatrix}
    x \\
    y
  \end{bmatrix}.
\end{align} 
The parameters $\omega$ and $\alpha$ controls the scaling and rotation, respectively, of the major and minor axes of a level curve of equal correlation (Figure$~$\ref{fig-tropy}). Using these transformed coordinates, the partial sill ($\sigma^2_{de}$), nugget ($\sigma^2_{ie}$), and range ($\phi$) parameters are estimated. We accommodate geometric anisotropy by supplying \code{anisotropy}:
```{r}
sppois_anis <- update(sppois, anisotropy = TRUE)
spnb_anis <- update(spnb, anisotropy = TRUE)
```

According to BIC, the spatial negative binomial model with geometric anisotropy performs best:
```{r}
BIC(sppois, spnb, sppois_anis, spnb_anis)
```

The \code{plot()} function can be used to visualize the anisotropy (Figure$~$\ref{fig-tropy}):
```{r, eval = FALSE}
plot(spnb, which = 8)
plot(spnb_anis, which = 8)
```

The spatial correlation is strongest in a northwest-southeast direction and weakest in the northeast-southwest direction (Figure$~$\ref{fig-tropy}), which is intuitive given the similar patterns in moose counts from Figure$~$\ref{fig-moose-data-count}.

\begin{figure}[h]
\centering
\includegraphics[width = 1\linewidth]{figures/figure-7.png}
\caption{Level curves of equal correlation for the negative binomial moose count models. The ellipse is centered at zero distance in the x-direction and y-direction, and points along the ellipse have equal levels of correlation.  In the isotropic level curve (left), spatial covariance decays equally in all directions. In the anistropic level curve (right), spatial covariance decays fastest in the northeast-southwest direction and slowest in the northwest-southeast direction (this pattern can be seen in the observed counts).}
\label{fig-tropy}
\end{figure}

## Modeling lake conductivity in Southwest, USA {#sec-lake}

The \code{lake} data in \code{spmodel} contains climate and chemical data for several lakes in four southwestern states in the United States: Arizona, Colorado, Nevada, and Utah. We desire an SPGLM that characterizes the effect of temperature, state, and lake origin (whether the lake is naturally occurring or human made) on lake conductivity. Conductivity is a measure of dissolved ions (measured here in water), which is important for various physical, chemical, and biological processes. Chemical data is often heavily right-skewed, so we model it using an SPGLM assuming a Gamma distribution for the response. The \code{log_cond} variable in \code{lake} is the logarithm of conductivity, which we dynamically exponentiate within \code{formula} so that it is on the original scale:
```{r}
spgam <- spglm(
  formula = exp(log_cond) ~ temp + state + temp:state + origin, 
  family = "Gamma",
  data = lake,
  spcov_type = "cauchy",
  partition_factor = ~ year
)
```

We model conductivity as a function of temperature, state, and lake origin, and we allow the effect of temperature to vary by state (\code{temp:state}). The \code{year} partition factor (specified via \code{partition_factor}) restricts spatial covariance to be nonzero only for observations sampled during the same year. Data were collected in 2012 and 2017, so this partition factor assumes independence between observations in 2012 and 2017. While we used the partition factor here illustratively, more generally, the utility of partition factors can be highly context dependent.

When categorical variables have more than two levels, the default reference group contrasts are not well-suited to assess the variable's overall significance:
```{r}
summary(spgam)
```

A more effective approach is to use an analysis of variance (ANOVA), which is well-suited to asses the overall significance of each variable:
```{r}
anova(spgam)
```

The main effect for temperature and the temperature by state interaction are highly significant ($p$-value < 0.001), while the main effects for state and lake origin are not significant. The ANOVA table can be tidied using \code{tidy()}.

Variance inflation factors assess the degree to which standard errors $\hat{\boldsymbol{\beta}}$ are inflated due to covariance among the columns of $\mathbf{X}$. Generalized variance inflation factors can capture the variance inflation for subsets of $\mathbf{X}$ that may include categorical variables with more than two levels [@fox1992generalized]:

```{r, warning=FALSE, message=FALSE}
library(car)
```

```{r}
vif(spgam)
```

The GVIF$^{1/2df}$ values for \code{temp}, \code{state}, and \code{temp:state} are just greater than two, which suggests moderate multicollinearity for these terms -- unsurprising given the \code{temp:state} interaction in the model. The GVIF$^{1/2df}$ for \code{origin} is close to one, which suggests little to no multicollinearity for this term.

Because of the interaction between \code{temp} and \code{state}, contrasts that assess mean differences among states should condition upon a specific temperature value. By default, \pkg{emmeans} uses the mean temperature value (here, 7.63) to assess contrasts:
```{r, warning=FALSE, message=FALSE}
library(emmeans)
```

```{r}
pairs(emmeans(spgam, ~ state | temp))
```

Again, because of the interaction between \code{temp} and \code{state}, we should assess temperature trends separately for each state:
```{r}
emtrends(spgam, ~ state, var = "temp")
```

## Modeling harbor seal trends in Alaska, USA {#sec-seal}

The \code{seal} data in \pkg{spmodel} contains harbor seal abundance trends for two different harbor seal stocks (genetically distinct populations). While the \code{moose} and \code{lake} data were point-referenced, the \code{seal} data are areal. Each polygon in the \code{seal} data represents a distinct harbor seal haulout location (Figure$~$\ref{fig-seal}). A haulout location is a spot on coastal rocks that harbor seals go to rest, molt, and give birth.

\begin{figure}
\centering
\includegraphics[width = 1\linewidth]{figures/figure-8.png}
\caption{Seal trend distribution in Alaska. Observed and missing seal polygons by stock (left) and observed log seal trends (right).}
\label{fig-seal}
\end{figure}

For each polygon, a Poisson regression was used to quantify the mean trend in abundance over approximately 30 years [@ver2018spatial]. If the logarithm of mean abundance trends (\code{log_trend}) is negative (positive), it means abundance is decreasing (increasing). We use a binomial SPGLM to quantify the likelihood that mean abundance trends are decreasing:
```{r}
is_decreasing <- seal$log_trend < 0
spbin <- spgautor(
  formula = is_decreasing ~ 1,
  family = binomial,
  data = seal,
  spcov_type = "car",
  random = ~ stock
)
```

To model spatial dependence, we used a conditional autoregressive function. Conditional and simultaneous autoregressive functions characterize spatial distance through neighborhood relationships (rather than Euclidean distance) and have \code{spcov_type} values of \code{"car"} and \code{"sar"}, respectively. By default, Queen's distance is used to determine whether two sites are neighbors, though custom neighborhood matrices are passed via \code{W}. Row standardization is also assumed by default; this can be changed via \code{row_st}. Using \code{random}, we also specified a nonspatial random effect for seal stock. The \code{random} argument uses similar syntax as \pkg{lme4} [@bates2015lme4] and \pkg{nlme} [@pinheiro2006mixed] to specify nonspatial random effects.

Tidying the model reveals the estimates and confidence intervals on the log odds scale:
```{r}
tidy(spbin, conf.int = TRUE)
```

Back-transforming the confidence interval to the probability scale yields:
```{r}
emmeans(spbin, ~ 1, type = "response")
```

The \code{SE} column is the standard error on the response scale obtained from the delta method [@oehlert1992note].

In contrast to point-referenced data, prediction locations for areal data must be specified at the time of model fitting, as they affect the spatial covariance function's neighborhood structure. Prediction locations whose response values have an \code{NA} (i.e., missing) value are converted into a \code{newdata} object that is stored in the model output. For example, rows one and nine are locations without seal trends, meaning they are not used in model fitting but are desired for prediction:
```{r}
seal
```

Then, \code{predict()} can be called without having to specify \code{newdata}:

```{r}
predict(spbin, type = "response", interval = "prediction")[1:5, ]
```

We could have alternatively used a (geostatistical) SPGLM via \code{spglm()}. When areal data are used with \code{spglm()}, the centroids of each polygon are used as the point-referenced coordinates. We further explore comparisons between point-referenced and aeral data in the next example.

## Modeling voter turnout in Texas, USA {#sec-texas}

\begin{figure}
\centering
\includegraphics[width = 0.70\linewidth]{figures/figure-10.png}
\caption{Proportion of voter turnout in Texas for the 1980 presidential election. Circles represent voter turnout (based on color) and triangles represent locations at which voter turnout predictions are desired.}
\label{fig-texas}
\end{figure}

The \code{texas} data in \pkg{spmodel} contains voter turnout data for Texas counties in the 1980 United States Presidential Election [@bivand2024spdata]. The data are point-referenced, with polygon centroids representing the spatial location of each county (Figure$~$\ref{fig-texas}). Beta regression is a GLM used to model rate and proportion data in the (0, 1) interval [@ferrari2004beta; @cribari2010beta]. We model voter turnout rates as a function of mean log income of county residents using an SPGLM assuming a beta distributed response variable:
```{r}
spbeta_geo <- spglm(
  formula = turnout ~ log_income,
  family = "beta", 
  data = texas,
  spcov_type = "matern"
)
```

Alternatively, we could use an autoregressive model to fit the model, constructing a neighborhood matrix by assuming centroids within \code{cutoff} of one another are neighbors:
```{r}
spbeta_auto <- spgautor(
  formula = turnout ~ log_income,
  family = "beta", 
  data = texas,
  spcov_type = "car",
  cutoff = 1e5
)
```

According to AIC, the SPGLM for point-referenced data is preferred:
```{r}
AIC(spbeta_geo, spbeta_auto)
```

The default estimation method in \pkg{spmodel} is restricted maximum likelihood (REML). One drawback of REML is that likelihood-based statistics are only valid for model comparison when the models have the same explanatory variable and fixed effect structure. This is because the error contrasts used to construct the REML likelihood change based on $\mathbf{X}$ and $\boldsymbol{\beta}$. An alternative is to use maximum likelihood, which can use likelihood-based statistics to compare models with different explanatory variable and fixed effect structures. While we could assess the significance of log turnout with the \code{spbeta_geo} model fit with REML (e.g., via \code{tidy()}), an alternative approach is to use a likelihood ratio test:
```{r}
spbeta_full_ml <- update(spbeta_geo, estmethod = "ml")
spbeta_red_ml <- update(spbeta_geo, estmethod = "ml", formula = turnout ~ 1)
anova(spbeta_full_ml, spbeta_red_ml)
```

The likelihood ratio test suggests that log income is significantly related to voter turnout ($p$-value < 0.001). We could instead use a different likelihood-based statistic like AIC:
```{r}
AIC(spbeta_full_ml, spbeta_red_ml)
```

The AIC prefers the full model, which suggests that log income is important for predicting voter turnout.

# Discussion {#sec-discussion}

SPGLMs are fit in \pkg{spmodel} using a novel application of the Laplace approximation that simultaneously marginalizes over the latent (i.e., unobserved) random effects and the fixed effects, $\boldsymbol{\beta}$. The approach is very flexible and accommodates general response distributions and covariance structures, though here we focus on spatial applications.  \pkg{spmodel}'s \code{spglm()} and \code{spgautor()} fit SPGLMs that are similar in structure and syntax as base \proglang{R}'s \code{glm()} function, easing the transition from GLMs to SPGLMs for practitioners. The \code{spglm()} and \code{spgautor()} functions support six response distributions for binary, count, and skewed data and 20 spatial covariance functions. \pkg{spmodel} also provides a suite of tools for data visualization, inference, model diagnostics, and prediction, providing a framework that can be used for all stages of a data analysis. There are many additional \pkg{spmodel} features that are not covered here, including fitting multiple models simultaneously, fixing spatial covariance and dispersion parameters at known values, fitting models to large non-Gaussian data having thousands of observations via spatial indexing [@ver2023indexing], incorporating spatial dependence in machine learning (e.g., random forests; @breiman2001random), simulating spatially dependent data (e.g., \code{spbinom()}, \code{sprpois()}, etc.), and more. Further details are provided by [https://CRAN.R-project.org/package=spmodel](https://CRAN.R-project.org/package=spmodel) and links therein.

# Computational details {.unnumbered}

The results in this paper were obtained using [R]{.proglang} 4.4.0 with the
\pkg{spmodel} 0.11.0 package. Figures were created using the [ggplot2]{.pkg} 3.5.1 package [@wickham2016ggplot2] and base [R]{.proglang}.

# Data and code availability {.unnumbered}

All writing and code associated with this manuscript is available for viewing and download on GitHub at [https://github.com/USEPA/spmodel.glm.manuscript](https://github.com/USEPA/spmodel.glm.manuscript). All data used are part of the \pkg{spmodel} [R]{.proglang} package available for download from CRAN at [https://CRAN.R-project.org/package=spmodel](https://CRAN.R-project.org/package=spmodel).

# Acknowledgments {.unnumbered}

We would like to thank initial reviewers and editors for feedback that has greatly improved the manuscript.

The views expressed in this manuscript are those of the authors and do not necessarily represent the views or policies of the U.S. Environmental Protection Agency or the National Oceanic and Atmospheric Administration. Any mention of trade names, products, or services does not imply
an endorsement by the U.S. government, the U.S. Environmental Protection Agency, or the National Oceanic and Atmospheric
Administration. The U.S. Environmental Protection Agency and the National Oceanic and Atmospheric Administration do not endorse
any commercial products, services or enterprises.

\bibliography{references.bib}